A New Method of Modelling Tuneable Lasers with Functional Composition

### 2019-07-12 Thesis/Appendices/Asymptotics.tex
Finished final read through.

% !TeX root = ../Thesis.tex
%\chapter{Asymptotic Expansions of the Variance}
%\label{chap:asymp}
We shall investigate the nature of the solution to \eqref{eq:var} both when $s \rightarrow 0$ (a very short \gls{cfbg}), as well as when $s \rightarrow \infty$ (a very long \gls{cfbg}). For ease of notation, \eqref{eq:var} is rewritten as
\begin{align}
\label{eq:asymp}
\frac{1}{4} x^4 = - \eps (x - 1)(x - 2)^2,
\end{align}
where $x = \sigma^2$, $\eps = s^4$, and $\eps \rightarrow 0$. Figure \ref{fig:asympt} shows \eqref{eq:asymp}. Additionally, suppose $x$ can be expanded as a power series in $\eps$:
\begin{align}
\label{eq:series}
x = x_0 + x_1 \eps^\alpha + x_2 \eps^\beta + \cdots
\end{align}
with $0 < \alpha < \beta$. Then, expanding \eqref{eq:asymp} using \eqref{eq:series} and collecting terms, at $\bigO{1}$,
\begin{align*}
\frac{1}{4} x_0^4 = 0,
\end{align*}
and so, $x_0 = 0$. Knowing this, we consider how to balance
\begin{align*}
\frac{1}{4} x_1^4 \eps^{4\alpha} = \eps \left( -x_1^3 \eps^{3\alpha} + 5x_1^2 \eps^{2\alpha} - 8x_1 \eps^\alpha + 4 \right).
\end{align*}
In order to have dominant balance it must be that the left hand side balances with the final term of the right hand side, that is, $4\alpha = 1$, or $\alpha = \frac{1}{4}$. Therefore, at $\bigO{\eps}$ $x_1^4 = 16$, or
\begin{align*}
(x_1 - 2)(x_1 + 2)(x_1^2 + 4) &= 0,
\end{align*}
and so, $x_1 = \pm 2, \, \pm 2 i$. However, we require $x$ to be positive and real---recall that $x = \sigma^2$---thus, we take $x_1 = 2$. Now, the next lowest order of the left hand side must be $\eps^{3\alpha + \beta} = \eps^{3/4 + \beta}$. As with the previous iteration, this must balance with the lowest order term of the $8x$. Hence, $\frac{3}{4} + \beta = 1 + \alpha = \frac{5}{4}$, thus, $\beta = \frac{1}{2}$. Now, at $\bigO{\eps^{5/4}}$,
\begin{align*}
\frac{1}{4} 4 x_1^3 x_2 &= -8x_1,
\end{align*}
and with the choice of $x_1$, $x_2 = 2$. Combining these results gives $x = 2 \eps^{1/4} - 2 \eps^{1/2} + \bigO{\eps^{3/4}}$, or $\sigma^2 = 2 s \left( 1 - s \right) + \bigO{s^3}$,
this is shown in Figure \ref{fig:lims0}. The choice $\eps = s^4$ was for convenience, but in retrospect, $\eps = s$ yields a cleaner, but identical, solution. \\
\begin{figure}[tbp]
\input{./Figures/Lim_s_0}
\caption[Asymptotic expansion of the variance as the dispersion parameter approaches 0.]{Asymptotic expansion of the variance as the dispersion parameter, $s$, approaches $0$.}
\label{fig:lims0}
\end{figure}

We shall now consider the other limit, when the dispersion parameter, $s$, approaches $\infty$. Using a similar substitution we instead write \eqref{eq:var} as
\begin{align}
\label{eq:asymp2}
\frac{1}{4} \eps x^4 = -x^3 + 5x^2 - 8x + 4,
\end{align}
where $x = \sigma^2$, and $\eps = s^{-4}$ instead so we still have $\eps \rightarrow 0$, this is also shown in Figure~\ref{fig:asympt}. Furthermore, we must make a correction to the series expansion:
\begin{align}
\label{eq:series2}
x = \eps^{-\xi} \left( x_0 + x_1 \eps^\alpha + x_2 \eps^\beta + \cdots \right).
\end{align}
The reason for this is because the equation is now singular---when $\eps=0$ the equation transforms from a quartic to a cubic, losing a root. As with before, we obtain dominant balance when the quartic term balances with the cubic term which occurs when $1 - 4 \xi = -3 \xi$, or $\xi = 1$. Then to $\bigO{\eps^{-3}}$, expression \eqref{eq:asymp2} gives
\begin{align*}
x_0^3 (x_0 + 4) &= 0.
\end{align*}
The non-trivial solution $x_0 = -4$ gives rise to a boundary layer structure. However, this is unphysical because $x \geq 0$, leaving $x_0^3 = 0$, so this singular perturbation does not generate the boundary layer expected. Therefore, \eqref{eq:series2} reduces to a regular perturbation problem, and $\alpha$ must equal 1. The next condition occurs at $\bigO{1}$ where
\begin{align*}
0 = -x_1^3 + 5x_1^2 - 8x_1 + 4 = -(x_1 - 1)(x_1 - 2)^2,
\end{align*}
so either $x_1 = 1$ or $x_1 = 2$. The next term obtains dominant balance when the left hand side balances with the linear term of the right hand side, that is, when $\beta = 2$ provided\footnote{In the case of $x_1 = 2$ we find $\beta = \frac{3}{2}$ which leads to complex roots.} $x_1 \ne 2$. Now at $\bigO{\eps}$,
\begin{align*}
\frac{1}{4} x_1^4 = -3x_1^2 x_2 + 5\cdot2 x_1 x_2 - 8x_2 = -x_2 (3x_1 - 4)(x_1 - 2),
\end{align*}
where again, the coefficients come from the binomial expansions. From this it is clear that
\begin{align*}
x_2 &= \frac{-x_1^4}{4(3x_1 - 4)(x_1 - 2)}.
\end{align*}
Thus, $x_1 = 1$, $x_2 = -\frac{1}{4}$, and
\begin{align*}
x = 1 - \frac{1}{4} \eps + x_2 \eps^{\gamma - 1} + \cdots.
\end{align*}
Expanding one final term we find $\gamma = 3$, and to $\bigO{\eps^2}$,
\begin{align*}
\frac{1}{4} 4x_2 x_1^3 &= -(3x_3 x_1^2 + 3x_2^2 x_1) + 5(2x_3 x_1 + x_2^2) - 8x_3.
\end{align*}
Knowing the values of $x_1$, and $x_2$ this is simply an arithmetical calculation yielding $x_3 = \frac{3}{8}$. Finally,
\begin{align*}
x &= 1 - \frac{1}{4}\eps + \frac{3}{8}\eps^2 + \bigO{\eps^3},
\end{align*}
or,
\begin{align*}
\sigma^2 &= 1 - \frac{1}{4s^4} + \frac{3}{8s^8} + \bigO{s^{-12}}
\end{align*}
this approximation is shown in Figure \ref{fig:limsinfty}. \\
\begin{figure}[tbp]
\input{./Figures/Lim_s_Infty}
\caption[Asymptotic expansion of the variance as the dispersion parameter approaches $\infty$.]{Asymptotic expansion of the variance as the dispersion parameter, $s$, approaches $\infty$.}
\label{fig:limsinfty}
\end{figure}

>>>>

% !TeX root = ../Thesis.tex
%\chapter{Asymptotic Expansions of the Variance}
%\label{chap:asymp}
We shall investigate the nature of the solution to \eqref{eq:var} both when $s \rightarrow 0$ (a very short \gls{cfbg}), as well as when $s \rightarrow \infty$ (a very long \gls{cfbg}). For ease of notation, \eqref{eq:var} is rewritten as
\begin{align}
\label{eq:asymp}
\frac{1}{4} x^4 = - \eps (x - 1)(x - 2)^2,
\end{align}
where $x = \sigma^2$, $\eps = s^4$, and $\eps \rightarrow 0$. Figure \ref{fig:asympt} shows \eqref{eq:asymp} graphically. Additionally, suppose $x$ can be expanded as a power series in $\eps$:
\begin{align}
\label{eq:series}
x = x_0 + x_1 \eps^\alpha + x_2 \eps^\beta + \cdots
\end{align}
with $0 < \alpha < \beta$. Then, expanding \eqref{eq:asymp} using \eqref{eq:series} and collecting terms, at $\bigO{1}$,
\begin{align*}
\frac{1}{4} x_0^4 = 0,
\end{align*}
and so, $x_0 = 0$. Knowing this, we consider how to balance
\begin{align*}
\frac{1}{4} x_1^4 \eps^{4\alpha} = \eps \left( -x_1^3 \eps^{3\alpha} + 5x_1^2 \eps^{2\alpha} - 8x_1 \eps^\alpha + 4 \right).
\end{align*}
In order to have dominant balance it must be that the left hand side balances with the final term of the right hand side, that is, $4\alpha = 1$, or $\alpha = \frac{1}{4}$. Therefore, at $\bigO{\eps}$ $x_1^4 = 16$, or
\begin{align*}
(x_1 - 2)(x_1 + 2)(x_1^2 + 4) &= 0,
\end{align*}
and so, $x_1 = \pm 2, \, \pm 2 i$. However, we require $x$ to be positive and real---recall that $x = \sigma^2$---thus, we take $x_1 = 2$. Now, the next lowest order of the left hand side must be $\eps^{3\alpha + \beta} = \eps^{3/4 + \beta}$. As with the previous iteration, this must balance with the lowest order term of the $8x$. Hence, $\frac{3}{4} + \beta = 1 + \alpha = \frac{5}{4}$, thus, $\beta = \frac{1}{2}$. Now, at $\bigO{\eps^{5/4}}$,
\begin{align*}
\frac{1}{4} 4 x_1^3 x_2 &= -8x_1,
\end{align*}
and with the choice of $x_1$, $x_2 = 2$. Combining these results gives $x = 2 \eps^{1/4} - 2 \eps^{1/2} + \bigO{\eps^{3/4}}$, or $\sigma^2 = 2 s \left( 1 - s \right) + \bigO{s^3}$,
this is shown in Figure \ref{fig:lims0}. The choice $\eps = s^4$ was for convenience, but in retrospect, $\eps = s$ yields a cleaner, but identical, solution. \\
\begin{figure}[tbp]
\input{./Figures/Lim_s_0}
\caption[Asymptotic expansion of the variance as the dispersion parameter approaches 0.]{Asymptotic expansion of the variance as the dispersion parameter, $s$, approaches $0$.}
\label{fig:lims0}
\end{figure}

We shall now consider the other limit, when the dispersion parameter, $s$, approaches $\infty$. Using a similar substitution we instead write \eqref{eq:var} as
\begin{align}
\label{eq:asymp2}
\frac{1}{4} \eps x^4 = -x^3 + 5x^2 - 8x + 4,
\end{align}
where $x = \sigma^2$, and $\eps = s^{-4}$ instead so we still have $\eps \rightarrow 0$, this is also shown in Figure~\ref{fig:asympt}. Furthermore, we must make a correction to the series expansion:
\begin{align}
\label{eq:series2}
x = \eps^{-\xi} \left( x_0 + x_1 \eps^\alpha + x_2 \eps^\beta + \cdots \right).
\end{align}
The reason for this is because \eqref{eq:asymp2} is now singular---when $\eps=0$ the equation transforms from a quartic to a cubic, losing a root. As with before, we obtain dominant balance when the quartic term balances with the cubic term which occurs when $1 - 4 \xi = -3 \xi$, or $\xi = 1$. Then to $\bigO{\eps^{-3}}$, expression \eqref{eq:asymp2} gives
\begin{align*}
x_0^3 (x_0 + 4) &= 0.
\end{align*}
The non-trivial solution $x_0 = -4$ gives rise to a boundary layer structure. However, this is unphysical because $x \geq 0$, leaving $x_0^3 = 0$, so this singular perturbation does not generate the boundary layer expected. Therefore, \eqref{eq:series2} reduces to a regular perturbation problem, and $\alpha$ must equal 1. The next condition occurs at $\bigO{1}$ where
\begin{align*}
0 = -x_1^3 + 5x_1^2 - 8x_1 + 4 = -(x_1 - 1)(x_1 - 2)^2,
\end{align*}
so either $x_1 = 1$ or $x_1 = 2$. The following term obtains dominant balance when the left hand side balances with the linear term of the right hand side, that is, when $\beta = 2$ provided\footnote{In the case of $x_1 = 2$ we find $\beta = \frac{3}{2}$ which leads to complex roots.} $x_1 \ne 2$. Now at $\bigO{\eps}$,
\begin{align*}
\frac{1}{4} x_1^4 = -3x_1^2 x_2 + 5\cdot2 x_1 x_2 - 8x_2 = -x_2 (3x_1 - 4)(x_1 - 2),
\end{align*}
where the coefficients come from the binomial expansions. From this it is clear that
\begin{align*}
x_2 &= \frac{-x_1^4}{4(3x_1 - 4)(x_1 - 2)}.
\end{align*}
Thus, $x_1 = 1$, $x_2 = -\frac{1}{4}$, and
\begin{align*}
x = 1 - \frac{1}{4} \eps + x_2 \eps^{\gamma - 1} + \cdots.
\end{align*}
Expanding one final term we find $\gamma = 3$, and to $\bigO{\eps^2}$,
\begin{align*}
\frac{1}{4} 4x_2 x_1^3 &= -(3x_3 x_1^2 + 3x_2^2 x_1) + 5(2x_3 x_1 + x_2^2) - 8x_3.
\end{align*}
Knowing the values of $x_1$ and $x_2$ this is simply an arithmetical calculation yielding $x_3 = \frac{3}{8}$. Finally,
\begin{align*}
x &= 1 - \frac{1}{4}\eps + \frac{3}{8}\eps^2 + \bigO{\eps^3},
\end{align*}
or,
\begin{align*}
\sigma^2 &= 1 - \frac{1}{4s^4} + \frac{3}{8s^8} + \bigO{s^{-12}}
\end{align*}
this approximation is shown in Figure \ref{fig:limsinfty}. \\
\begin{figure}[tbp]
\input{./Figures/Lim_s_Infty}
\caption[Asymptotic expansion of the variance as the dispersion parameter approaches $\infty$.]{Asymptotic expansion of the variance as the dispersion parameter, $s$, approaches $\infty$.}
\label{fig:limsinfty}
\end{figure}

### 2019-07-12 Thesis/Appendices/Hermite.tex
Finished final read through.

\begin{theorem}
\label{thm:gaussian}
Gaussians of a single variance span $L^2(\mathbb{R})$.
\end{theorem}
\begin{proof}
Theorem \ref{thm:hermite} showed that any square integrable function can be expressed as
\begin{align*}
f(x) = \sum_{n = 0}^\infty a_n \Her{n}{x}.
\end{align*}
Recall that $\Her{n}{x}$ has a polynomial as well as a Gaussian component. Without loss of generality we assumed the Gaussian has variance $\frac{1}{2}$ for convenience. We can effectively remove the polynomial dependence by using a similar idea to \cite{calcaterra2, calcaterra}. We shall expand the Gaussian--Hermite polynomials using their definition
\begin{align*}
f(x) = \sum_{n = 0}^\infty a_n (-1)^n \textrm{e}^{x^2/2} \diff[n]{}{x} \textrm{e}^{-x^2}.
\end{align*}
Finally, we can rewrite the derivatives using central differences \cite{zwillinger} so that
\begin{align}
f(x) = \sum_{n = 0}^\infty a_n (-1)^n \textrm{e}^{x^2/2} \left[ \lim_{h \rightarrow 0} \frac{1}{h^n} \sum_{i = 0}^n (-1)^i \binom{n}{i} \exp\left( -\left( x + h\left( \frac{n}{2} - i \right) \right)^2 \right) \right].
\label{eq:gs}
\end{align}
\end{proof}

@@

Two examples of Gaussian series from \eqref{eq:gs} are shown in Figure \ref{fig:gs}. These examples are taken from \cite{calcaterra2, calcaterra}, however, we achieve similar or better approximations with either fewer terms, or with an $h$ value an order of magnitude larger. There are two reasons for this, first, we use central difference as opposed to backwards difference yielding a convergence of $\bigO{h^2}$ instead of $\bigO{h}$. Furthermore, because \cite{calcaterra2, calcaterra} used backwards differences, the means of the Gaussians are all non-negative, whereas with central differences, our Gaussians' means are both positive and negative---this leads to smaller coefficients, and better numerical stability. Moreover, Figure \ref{fig:gaussmod} shows the Gaussian series for the modulation function in \cite{bohun, burgoyneemail}.

>>>>

\begin{theorem}
\label{thm:gaussian}
Gaussians of a single variance span $L^2(\mathbb{R})$.
\end{theorem}
\begin{proof}
Theorem \ref{thm:hermite} showed that any square integrable function can be expressed as
\begin{align*}
f(x) = \sum_{n = 0}^\infty a_n \Her{n}{x}.
\end{align*}
Recall that $\Her{n}{x}$ has a polynomial as well as a Gaussian component. Without loss of generality we assumed the Gaussian has variance $\frac{1}{2}$ for convenience. We can effectively remove the polynomial dependence by using a similar idea to \cite{calcaterra2, calcaterra}. We shall expand the Gaussian--Hermite polynomials using their definition
\begin{align*}
f(x) = \sum_{n = 0}^\infty a_n (-1)^n \textrm{e}^{x^2/2} \diff[n]{}{x} \textrm{e}^{-x^2}.
\end{align*}
Finally, we can rewrite the derivatives using central differences \cite{zwillinger} so that our function can be expressed as
\begin{align}
f(x) = \sum_{n = 0}^\infty a_n (-1)^n \textrm{e}^{x^2/2} \left[ \lim_{h \rightarrow 0} \frac{1}{h^n} \sum_{i = 0}^n (-1)^i \binom{n}{i} \exp\left( -\left( x + h\left( \frac{n}{2} - i \right) \right)^2 \right) \right].
\label{eq:gs}
\end{align}
\end{proof}

@@

Two examples of Gaussian series from \eqref{eq:gs} are shown in Figure \ref{fig:gs}. These examples are taken from \cite{calcaterra2, calcaterra}, however, we achieve similar or better approximations with either fewer terms, or with an $h$ value an order of magnitude larger. There are two reasons for this, first, we use central difference as opposed to backwards difference yielding a convergence of $\bigO{h^2}$ instead of $\bigO{h}$. Furthermore, because \cite{calcaterra2, calcaterra} used backwards differences, the means of the Gaussians are all non-negative, whereas with central differences, our Gaussians' means are both positive and negative---this leads to smaller coefficients, and better numerical stability. Moreover, Figure \ref{fig:gaussmod} shows the Gaussian series for the modulation function
\begin{align*}
M(T) = \frac{1}{2} \left( 1 - \cos \left( \pi \textrm{e}^{-x^2} \right) \right),
\end{align*}
used in \cite{bohun, burgoyneemail}.

### 2019-07-12 Thesis/Conclusion/Conclusion.tex
Finished final read through.

One of the issues experimentalists have is designing their tuneable lasers. They are uncertain how their laser will behave---and if the laser will lase at all---until they actually construct it. This model serves to provide better design principles, and how to troubleshoot should they encounter wave breaking. For example, from our results it is clear that to fix a broken wave $b$ must be decreased, or $s$ increased. This could be achieved by reducing the length of the optical fibre between the gain fibre and the output coupler, or by increasing $\beta_2 L_D$ by adding a second \gls{cfbg} in series\footnote{See \eqref{eq:ndparam}.}. Additionally, one could decrease  the modulation time, $T_M$, to move through the parameter space along a ray outward from the origin. Although this would increase $b$, it also increases $s$ enough to move out of the unstable region. \\

>>>>

One of the issues experimentalists have is designing their tuneable lasers. They are uncertain how their laser will behave---and if the laser will lase at all---until they actually construct it. This model serves to provide better design principles, and how to troubleshoot should they encounter wave breaking. For example, from our results it is clear that to fix a broken wave $b$ must be decreased, or $s$ increased. This could be achieved by reducing the length of the optical fibre between the gain fibre and the output coupler, or by increasing $\beta_2 L_D$ by adding a second \gls{cfbg} in series, respectively\footnote{See \eqref{eq:ndparam}.}. Additionally, one could decrease  the modulation time, $T_M$, to move through the parameter space along a ray outward from the origin. Although this would increase $b$, it also could increase $s$ enough to move out of the unstable region. \\

### 2019-07-12 Thesis/Intro/Introduction.tex
Finished final read through.

The word laser was originally an acronym for \emph{Light Amplification by Stimulated Emission of Radiation}. The cavity of a standard laser---such as a HeNe laser---is composed of four main components: a highly reflective mirror, a partially reflecting mirror, something to supply the energy, and the amplifying, or gain, medium. The two mirrors are at opposite ends of the cavity and serve to contain most of the light in the cavity while allowing a small amount to exit---this is the output of the laser, this is shown in Figure \ref{fig:laser}. The energy is generally supplied by either an external light source, or an electric field. The amplifying medium typically consists of gases, such as Helium and Neon, or a crystal, such as \gls{ndyag} for a single frequency laser. \\

\begin{figure}[tbp]
\centering
\input{./Figures/Emission}
\caption[Comparison of pumping, spontaneous emission, and stimulated emission.]{In pumping, incident photons are absorbed by electrons which are then excited into a much higher energy state (the electrons may also be excited by an electric field). The electrons quickly emits a low energy photon and transitions to an energy state that is still much more energetic than the original. In spontaneous emission, an electron in an excited state emits a photon and returns to the lower energy state. Finally, in stimulated emission, an incident photon interacts with an electron in an excited state, the electron donates its energy to the photon and returns to the lower energy state.}
\label{fig:emission}
\end{figure}

@@

Laser light has two fundamental characteristics that regular light does not. The first is that the light is highly monochromatic---the light contains (ideally) a single frequency. This is in contrast with an incandescent light bulb, for example, which emits light at all wavelengths with relative intensities given by Planck's law of black-body radiation. The other key feature is coherence, in which all the peaks and troughs of the light overlap giving very strong constructive interference---this is why laser light can be so intense. A system such as a \gls{led} does not produce laser light because, while monochromatic, the light is not coherent. The monochromatic nature is a result of how the light is generated; the frequency of the light emitted directly corresponds to the energy difference between the excited and ground states of the gain medium (as with \gls{led}s). Note that this means the frequency the laser operates at is fixed, and cannot be changed. Coherence is achieved in the laser cavity---the cavity is carefully designed so the light resonates with the length of the cavity. This causes all the peaks and troughs to align, and interfere constructively. \\

Another class of lasers are optical fibre lasers. Optical fibre lasers operate by the same principles, however, instead of the light being perpetually reflected back and forth, the light travels in a circular fashion through an optical fibre---this is a so called ring laser. Additionally, instead of the gain medium consisting of a gas or crystal, it is a length of fibre doped with a rare-earth metal. \\

@@

\subsection{Modulator}
The modulator serves the purpose of reshaping the pulse. Without it, the pulse will repeatedly widen due to the \gls{cfbg}, but the modulator ensures the pulse is band limited by altering the envelope. The optical pulse in the cavity can be modulated with one of two methods---amplitude modulation, or phase modulation. Mathematically these can be considered equivalent since they are Fourier transforms of each other, but in practice these are accomplished by different apparatuses. We shall assume the pulses are amplitude modulated to minimize the analysis required in the Fourier domain. Generally, the pulse can be amplitude modulated using one of two techniques. The first is with an acousto-optic modulator which utilizes the acousto-optic effect where the index of refraction of the optical fibre is varied by a sound wave \cite{hausbook, karim}. The second---and more common---method uses the electro-optic effect, where the index of refraction of the optical fibre is varied by a sound wave, and is called an electro-optic modulator. This modulator works the same was as an acousto-optic modulator, the optical fibre inside the modulator responds to an applied electric field, modifying the index of refraction \cite{agrawal2002, goldstein, hausbook, karim}. The shape of the electrical (sound) pulse can be controlled in such a way to provide the desired effect of modulation. \\

@@

\section{Mode-Locking}
\begin{figure}[tbp]
\centering
\begin{subfigure}{0.5\textwidth}
\input{./Figures/ML1}
\caption{$n = 4$, $\text{\acrshort{fwhm}} \approx 0.1255$}
\end{subfigure}%
\begin{subfigure}{0.5\textwidth}
\input{./Figures/ML2}
\caption{$n = 12$, $\text{\acrshort{fwhm}} \approx 0.04711$}
\end{subfigure}
\caption[Two examples of mode-locking.]{Two examples of mode-locking given by $\displaystyle \sum_{k = 1}^n \cos \left( 2 \pi k x \right)$.}
\label{fig:ml}
\end{figure}
One of the fundamental features of a laser is coherence. However, in the context of tuneable lasers it is perhaps less clear what is meant by coherence since the light is no longer monochromatic. Instead, we have a particular peak of all the frequencies aligned, that is, the phase shift between modes is zero. This process is called mode-locking---all of the different modes are locked together at a peak \cite{agrawal2013, hausbook, silfvast}. This of course creates a very intense and short pulse of light---as short as $5$ fs \cite{silfvast}. At this particular peak all of the frequencies constructively interfere as with a standard laser, however, deviations from the peak lead to destructive interference. Two examples of this are shown in Figure \ref{fig:ml}, and as expected, including more modes simultaneously narrows and intensifies the pulse. \\

>>>>

The word laser was originally an acronym for \emph{Light Amplification by Stimulated Emission of Radiation}. The cavity of a standard laser---such as a HeNe laser---is composed of four main components: a highly reflective mirror, a partially reflecting mirror, something to supply the energy, and the amplifying, or gain, medium. The two mirrors are at opposite ends of the cavity and serve to contain most of the light in the cavity while allowing a small amount to exit---this is the output of the laser, this is shown in Figure \ref{fig:laser}. The energy is generally supplied by either an external light source, or an electric field. The amplifying medium for a single frequency laser typically consists of gases, such as Helium and Neon, or a crystal, such as \gls{ndyag}. \\

\begin{figure}[tbp]
\centering
\input{./Figures/Emission}
\caption[Comparison of pumping, spontaneous emission, and stimulated emission.]{In pumping, incident photons are absorbed by electrons which are then excited into a much higher energy state (the electrons may also be excited by an electric field). The electrons quickly emit a low energy photon and transition down to an energy state that is still much more energetic than the original. In spontaneous emission, an electron in an excited state emits a photon and returns to the lower energy state. Finally, in stimulated emission, an incident photon interacts with an electron in an excited state, the electron donates its energy to the photon and returns to the lower energy state.}
\label{fig:emission}
\end{figure}

@@

Laser light has two fundamental characteristics that regular light does not. The first is that the light is highly monochromatic---the light contains (ideally) a single frequency. This is in contrast with an incandescent light bulb, for example, which emits light at all wavelengths with relative intensities given by Planck's law of black-body radiation. The other key feature is coherence, in which all the peaks and troughs of the light overlap giving very strong constructive interference---this is why laser light can be so intense. A system such as a \gls{led} does not produce laser light because, while monochromatic, the light is not coherent. The monochromatic nature is a result of how the light is generated; the frequency of the light emitted directly corresponds to the energy difference between the excited and ground states of the gain medium (as with \gls{led}s). Note that this means the operating frequency of the laser is fixed, and cannot be changed. Coherence is achieved in the laser cavity---the cavity is carefully designed so the light resonates with the length of the cavity. This causes all the peaks and troughs to align, and interfere constructively. \\

Another class of lasers are optical fibre lasers. Optical fibre lasers operate by the same principles, however, instead of the light being perpetually reflected back and forth, the light travels in a circular fashion along an optical fibre---this is a so called ring laser. Additionally, instead of the gain medium consisting of a gas or crystal, it is a length of fibre doped with a rare-earth metal. \\

@@

\subsection{Modulator}
The modulator serves the purpose of reshaping the pulse. Without it, the pulse will repeatedly widen due to the \gls{cfbg}, but the modulator ensures the pulse is band limited by altering the envelope. The optical pulse in the cavity can be modulated with one of two methods---amplitude modulation, or phase modulation. Mathematically these can be considered equivalent since they are Fourier transforms of each other, but in practice these are accomplished by different apparatuses. We shall assume the pulses are amplitude modulated to minimize the analysis required in the Fourier domain. Generally, the pulse can be amplitude modulated using one of two techniques. The first is with an acousto-optic modulator which utilizes the acousto-optic effect where the index of refraction of the optical fibre is varied by a sound wave \cite{hausbook, karim}. The second---and more common---method uses the electro-optic effect, where the index of refraction of the optical fibre is varied by an electric field, and is called an electro-optic modulator. This modulator works the same was as an acousto-optic modulator, the optical fibre inside the modulator responds to an applied electric field, modifying the index of refraction \cite{agrawal2002, goldstein, hausbook, karim}. The shape of the electrical (sound) pulse can be controlled in such a way to provide the desired effect of modulation. \\

@@

\section{Mode-Locking}
\begin{figure}[tbp]
\centering
\begin{subfigure}{0.5\textwidth}
\input{./Figures/ML1}
\caption{$n = 4$, $\text{\acrshort{fwhm}} \approx 0.1255$}
\end{subfigure}%
\begin{subfigure}{0.5\textwidth}
\input{./Figures/ML2}
\caption{$n = 12$, $\text{\acrshort{fwhm}} \approx 0.04711$}
\end{subfigure}
\caption[Two examples of mode-locking.]{Two examples of mode-locking given by $\displaystyle \sum_{k = 1}^n \cos \left( 2 \pi k x \right)$.}
\label{fig:ml}
\end{figure}
Recall, one of the fundamental features of a laser is coherence. However, in the context of tuneable lasers it is perhaps less clear what is meant by coherence since the light is no longer monochromatic. Instead, we have a particular peak of all the frequencies aligned, that is, the phase shift between modes is zero. This process is called mode-locking---all of the different modes are locked together at a peak \cite{agrawal2013, hausbook, silfvast}. This of course creates a very intense and short pulse of light---as short as $5$ fs \cite{silfvast}. At this particular peak all of the frequencies constructively interfere as with a standard laser, however, deviations from the peak lead to destructive interference. Two examples of this are shown in Figure \ref{fig:ml}, and as expected, including more modes simultaneously narrows and intensifies the pulse. \\
